\bf Calendar-based graphics for visualizing people's daily schedules

### 2018-10-23 blinded/blinded.tex
new directory blinded

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

\bigskip
\begin{abstract}
Calendars are broadly used in society to display temporal information,
and events. This paper describes a new R package with functionality to
organize and display temporal data, collected on sub-daily resolution,
into a calendar layout. The function \texttt{frame\_calendar} uses
linear algebra on the date variable to restructure data into a format
lending itself to calendar layouts. The user can apply the grammar of
graphics to create plots inside each calendar cell, and thus the
displays synchronize neatly with \textbf{ggplot2} graphics. The
motivating application is studying pedestrian behavior in Melbourne,
Australia, based on counts which are captured at hourly intervals by
sensors scattered around the city. Faceting by the usual features such
as day and month, was insufficient to examine the behavior. Making
displays on a monthly calendar format helps to understand pedestrian
patterns relative to events such as work days, weekends, holidays, and
special events. The layout algorithm has several format options and
variations. It is implemented in the R package \textbf{sugrrants}.
\end{abstract}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \textbf{ggplot2} \citep{R-ggplot2}. The data
restructuring approach is consistent with the tidy data principles
available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
methods are implemented in a new package called \textbf{sugrrants}
\citep{R-sugrrants}.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The
approach is developed with three conditions in mind: (1) to display
time-of-day variation in addition to longer temporal components such as
day-of-week and day-of-year; (2) to incorporate line graphs and other
types of glyphs into the graphical toolkit for the calendar layout; (3)
to enable overlaying plots consisting of multiple time series. The
proposed algorithm has been implemented in the \texttt{frame\_calendar}
function in the \textbf{sugrrants} package using R.

@@

Linking calendar displays to other types of charts is valuable to
visually explore the relationships between variables. An example can be
found in the \textbf{wanderer4melb} shiny application
\citep{R-wanderer4melb}. The calendar most naturally serves as a tool
for date selection: by selecting and brushing the glyphs in the
calendar, it subsequently highlights the elements of corresponding dates
in other time-based plots. Conversely, selecting on weather data plots,
linked to the calendar can help to assess if very hot/cold days and
heavy rain affect the number of people walking in downtown Melbourne.
The linking between weather data and calendar display is achieved using
the common dates.

@@

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

We would like to thank Stuart Lee and Heike Hofmann for their feedback
about this work. The most recent version of the \texttt{frame\_calendar}
function is included in the \textbf{sugrrants} package, which can be
accessed via the CRAN website
\url{https://CRAN.R-project.org/package=sugrrants} or Github
\url{https://github.com/earowang/sugrrants}. All materials required to
reproduce this article and a history of the changes can be found at the
project's Github repository
\url{https://github.com/earowang/paper-calendar-vis}.

>>>>

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

@@

\bigskip
\begin{abstract}
Calendars are broadly used in society to display temporal information,
and events. This paper describes a new R package with functionality to
organize and display temporal data, collected on sub-daily resolution,
into a calendar layout. The function \texttt{frame\_calendar} uses
linear algebra on the date variable to restructure data into a format
lending itself to calendar layouts. The user can apply the grammar of
graphics to create plots inside each calendar cell, and thus the
displays synchronize neatly with \textbf{ggplot2} graphics. The
motivating application is studying pedestrian behavior in Melbourne,
Australia, based on counts which are captured at hourly intervals by
sensors scattered around the city. Faceting by the usual features such
as day and month, was insufficient to examine the behavior. Making
displays on a monthly calendar format helps to understand pedestrian
patterns relative to events such as work days, weekends, holidays, and
special events. The layout algorithm has several format options and
variations.
\end{abstract}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \textbf{ggplot2} \citep{R-ggplot2}. The data
restructuring approach is consistent with the tidy data principles
available in the \textbf{tidyverse} \citep{R-tidyverse} suite. The
methods are implemented in a new package.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The
approach is developed with three conditions in mind: (1) to display
time-of-day variation in addition to longer temporal components such as
day-of-week and day-of-year; (2) to incorporate line graphs and other
types of glyphs into the graphical toolkit for the calendar layout; (3)
to enable overlaying plots consisting of multiple time series. The
proposed algorithm has been implemented in the \texttt{frame\_calendar}
function in a package using R.

@@

Linking calendar displays to other types of charts is valuable to
visually explore the relationships between variables. An example can be
found in the \textbf{wanderer4melb} shiny application. The calendar most
naturally serves as a tool for date selection: by selecting and brushing
the glyphs in the calendar, it subsequently highlights the elements of
corresponding dates in other time-based plots. Conversely, selecting on
weather data plots, linked to the calendar can help to assess if very
hot/cold days and heavy rain affect the number of people walking in
downtown Melbourne. The linking between weather data and calendar
display is achieved using the common dates.

@@
